Article Title

### 2013-12-03 plos.tex
Completed the  Results section.

We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (Fig.~\ref{fig:03}\,B and E) in the visual cortex on the contralateral side to the stimulus. With acousto-optic modulator (AOD) 3D steering of the laser, the located cells were imaged at high sampling rates with concurrent motion detection.  This technique allowed fast sampling (100-150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:03}\,C and D). Only cells that produced detectable calcium activity were included in the analysis. The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (Fig.~\ref{fig:03}\,F and G). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol for noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cells (Fig.~\ref{fig:03}\,E). Overall, 31 imaged sites from 24 animals were included in the study.

In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others. At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).

We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators. The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).

\hfill\rule{8cm}{4pt}

\subsection*{Relationship between functional covariance structure and circuit architecture}

\section*{Methods}
\subsection*{Estimation of parameters and hyperparameters of covariance estimators}

\subsection*{Regularization}
Under many loss functions\footnote{
The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. 
However, the equality does hold for its close cousin, Stein's \emph{entropy loss}, which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$. 
This deficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. 
The bias-variance decomposition is presented here to motivate the use of regularization.}, 
the risk can be decomposed as the sum
\begin{equation}\label{eq:bias-variance}
    r = b + \varepsilon
\end{equation}
of \emph{approximation error} (``bias'' or systematic error)
\begin{equation}
   b = \loss{\bar\Sigma,\Sigma}
\end{equation}
and \emph{estimation error} (``variance'') 
\begin{equation}
   \varepsilon = \E[\hat\Sigma]{\loss{\hat\Sigma, \bar\Sigma}}
\end{equation}
where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.

An unbiased estimator, such as the sample covariance,  $\hat\Sigma_0,$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.

The estimator risk can be reduced by \emph{regularization} -- a deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}. 
A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimate to minimize the estimator's risk.  
Various regularization schemes focus on the dimensionality reduction part  by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.  
Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}. 
Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.

\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
is then 
\begin{equation}\label{eq:sampleVariance}
\hat T= \hat\Sigma_0\circ I
\end{equation}
where $\circ$ denotes the entrywise matrix product (Hadamard product). 
Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
\begin{equation}\label{eq:equalVariance}
\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
\end{equation}
The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}. It is not immediately clear which target is better in a particular application. We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
\begin{equation}
\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
\end{equation}
The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
\end{equation}
This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.

We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by  cross-validation within the training dataset.

\subsubsection*{Estimate $\mathcal B$: Shrinkage toward a factor model}
\emph{Estimator $\mathcal B$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors, which corresponds to graphical models of the kind in Fig.~\ref{fig:02}B. For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by $d$ common inputs.

The target estimate is then 
\begin{equation}
\hat T_d = \hat L_d + \hat \Psi
\end{equation}
where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.

Just as with estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
\end{equation}
and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.

Factor analysis has long been used in natural sciences to infer the influence of a one or several latent variables exerting linear effects on a larger population of observed variables \cite{Jvreskog:1996}.  In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.  Estimation of covariance matrices by shrinkage toward a factor model has been used in finance for portfolio risk assessment \cite{Ledoit:2003,Fan:2008}.

\subsubsection*{Estimator $\mathcal C$: Spare inverse}
\emph{Estimator $\mathcal C$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse. This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972}.  Covariance matrices of this family correspond to graphical models of the type depicted in Fig.~\ref{fig:02}C.

For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statistical independence between the corresponding pairs. Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent. The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.

The covariance matrix estimate is then 
\begin{equation}
\hat\Sigma_\rho = \hat S^{-1}
\quad\mbox{with}\quad
\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
\end{equation}
where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
Solving this approximation in this form is computationally challenging. The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.  The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator  
\begin{equation}
\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
\quad\mbox{with}\quad
\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
\end{equation}
where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix.

Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.

\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix. Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.

The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.  With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.  When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
\begin{equation}
S^* = 
\begin{pmatrix}
S_{11} & S_{12} \\
S_{12}^\T & S_{22} 
\end{pmatrix}
\end{equation}
where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units. 
Then the covariance matrix of between the observed unit is 
\begin{equation}
\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation} 
The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units. Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.

Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
\begin{equation}
\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
\quad\mbox{where}\quad
(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
\end{equation}

Again, in this formulation, the approximation problem is computationally intractible.  The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm. The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
\begin{equation}
\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
\quad\mbox{where}\quad
(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
\end{equation}
The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}
Genevera Allen

>>>>

We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (Fig.~\ref{fig:03}\,B and E) in the visual cortex on the contralateral side to the stimulus. With acousto-optic modulator (AOD) 3D steering of the laser, the located cells were imaged at high sampling rates with concurrent motion detection.  This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:03}\,C and D). Only cells that produced detectable calcium activity were included in the analysis. The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (Fig.~\ref{fig:03}\,F and G). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol for noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cells (Fig.~\ref{fig:03}\,E). Overall, 31 imaged sites from 24 animals were included in the study.

In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others. At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).

We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators (Fig.~\ref{fig:04}). The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).

\subsection*{Relationship of correlation structure to circuit architecture}
Having demonstrated that $C_{\sf sparse+lowrank}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites. By its design, the sparse+lowrank estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

We examined the structure of the sparse+lowrank estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}. In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D). However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E). The partial correlation was decomposed into the sparse and low-rank components (Fig.\,\ref{fig:05}\,C). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), although mostly positive, had a much larger proportion of negative values. The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G). The low-rank component was of rank 17.

In previous publications that looked at the structure of correlations, the sparse network was obtained by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}. There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+lowrank estimator (Fig.\;\ref{fig:05}\,F). When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I). In particular, many low sample correlations translated into negative interactions in the regularized estimate. Indeed, the absence of a correlation between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient. Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.

The average partial correlations  revealed by the regularized estimator in all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (Fig.\;\ref{fig:06}\,A). The average node degree of the sparse component of the partial correlations and the number of  inferred latent units varied widely between sites but generally increased with the recorded population size (Fig.\;\ref{fig:06}\,B and C). However, there was an inverse relationship between the number of latent units and the average node degree (Fig.\;\ref{fig:06}\,D). Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.  These differences have multiple explanations, including differences in brain states and recording quality (neuropil contamination, motion).

We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells  to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}). The average partial correlations fell more rapidly with with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).

The connectivity of partial correlations in the regularized estimate had different organization for positive and negative interactions. Positive interactions fell rapidly as a function of difference in preferred orientation (Fig.\;\ref{fig:07}\,G and J), lateral displacements (Fig.\;\ref{fig:07}\,H and K), and displacement in depth (Fig.\;\ref{fig:07}\,I and L). Negative interactions were much less selective (Fig.\;\ref{fig:07}\,G--L).

\section*{Discussion}
TODO

\section*{Methods}
% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\subsection*{Data Acquisition}
Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
\subsection*{Covariance Estimation}
TODO
\subsection*{Simulation}
TODO
\subsection*{Cross-validation}}
TODO

\subsection*{Evaluation}
\section*{Acknowledgments}
Genevera Allen
